---
title: "Manchester City Transfer Options"
author: "Ben Irvine"
date: "22/09/2021"
output: ioslides_presentation
incremental: true
template: quarterly-report.html
widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(shiny)
library(shinythemes)
library(ggthemes)
```

## Brief
- Manchester City want to improve in the January transfer window.

- They don't want to add more than one player, as Pep Guardiola fears adding more could    upset the balance of his team.

- They want a proven talent as they can't afford anything but a top two finish this season and are not confident in doing so with their current squad.

- Where should they invest their money?

## Who are Manchester City?
- Football team founded in 1894

- Highly successful club competing in the Premier League

- Rapidly improved since being bought over by the Abu Dhabi United Group in 2008

- Only team to ever reach the 100 point mark in the Premier League

- Evaluated by Forbes at $2.69 billion

- Finished 1st in the Premier League just last year


## Approach to tackling the brief?
- Planned how I was going to meet their request

- Collected Appropriate data

- Cleaned the Data

- Looked at where Manchester City were statistically weakest

- Looked for Players to Strengthen their Weakness(ses)

## Preemptive Jargon Bust - What is xG?


xG (Expected goals) is a predictive model used to assess every goal-scoring chance, and the likelihood a player will score. An xG model computes, for each chance, the probability to score based on what we know about the outcome of similar chances.

- e.g a Penalty kick has an xG of 0.79

- a shot from the halfway line would have a very low xG

- It's very important that xG is interpreted correctly for it to have any kind of meaning


# Manchester City Strength and Weakness identification


---

```{r echo=FALSE,out.height="600px",out.width="1000px"}
team_xg_data <- read_csv("man_city_transfer_project/data/understat_per_game.csv") %>% 
  rename("conceded" = "missed")

xG_data_2019 <- team_xg_data %>%
 filter(league == "EPL",
        year == 2019) %>%
  group_by(team) %>%
  summarise(mean(xG))

xG_data_2019 %>%
  rename("Average_xG" = "mean(xG)") %>%
  ggplot() +
  aes(x = reorder(team, -Average_xG), y = Average_xG, fill = team) +
  geom_col(colour = "#000000") +
  theme_economist() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("#EF0107", "#95BFE5", "#DA291C", "#0057B8", "#6C1D45",
                               "#034694", "#1B458F", "#003399", "#003090", "#C8102E",
                               "#6CABDD", "#DA291C", "#241F20", "#FFF200", "#EE2737",
                               "#D71920", "#132257", "#FBEE23", "#7A263A", "#FDB913")) +
  labs(x = "\nTeam",
             y = "Average Expected Goals\n",
             colour = "",
             title = "Premier League Average xG per match\n")
```

---

```{r echo=FALSE,out.height="600px",out.width="1000px"}
team_xg_data %>%
 filter(league == "EPL",
        year == 2019) %>%
  group_by(team) %>%
  summarise(mean(xG_diff)) %>%
  rename("Average_xG_diff" = "mean(xG_diff)") %>%
  ggplot() +
  aes(x = reorder(team, -Average_xG_diff), y = Average_xG_diff, fill = team) +
  geom_col(colour = "#000000") +
  theme_economist() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("#EF0107", "#95BFE5", "#DA291C", "#0057B8", "#6C1D45",
                               "#034694", "#1B458F", "#003399", "#003090", "#C8102E",
                               "#6CABDD", "#DA291C", "#241F20", "#FFF200", "#EE2737",
                               "#D71920", "#132257", "#FBEE23", "#7A263A", "#FDB913")) +
  labs(x = "\nTeam",
             y = "Avg Diff Between xG and Goals \n",
             colour = "",
             title = "Average Difference Between xG and Goals Scored per match\n")
```

---

```{r echo=FALSE,out.height="600px",out.width="1000px"}
team_xg_data %>%
   filter(league == "EPL",
          year == 2019) %>%
  group_by(team) %>%
  summarise(mean(conceded)) %>%
   rename("Average_Goals_Conceded" = "mean(conceded)") %>%
  ggplot() +
  aes(x = reorder(team,  Average_Goals_Conceded), y = Average_Goals_Conceded, fill = team) +
  geom_col(colour = "#000000") +
  theme_economist() +
 theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("#EF0107", "#95BFE5", "#DA291C", "#0057B8", "#6C1D45",
                               "#034694", "#1B458F", "#003399", "#003090", "#C8102E",
                               "#6CABDD", "#DA291C", "#241F20", "#FFF200", "#EE2737",
                               "#D71920", "#132257", "#FBEE23", "#7A263A", "#FDB913")) +
  labs(x = "\nTeam",
             y = "Average Goals conceded\n",
             colour = "",
             title = "Premier League Avg Goals Conceded per Match\n")

```

---

```{r echo=FALSE,out.height="600px",out.width="1000px"}
team_xg_data %>%
   filter(league == "EPL",
          year == 2019) %>%
  group_by(team) %>%
  summarise(mean(xGA)) %>%
   rename("Average_Expected_Goals_Conceded" = "mean(xGA)") %>%
   ggplot() +
  aes(x = reorder(team, Average_Expected_Goals_Conceded), y = Average_Expected_Goals_Conceded, fill = team) +
  geom_col(colour = "#000000") +
  theme_economist() +
theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("#EF0107", "#95BFE5", "#DA291C", "#0057B8", "#6C1D45",
                               "#034694", "#1B458F", "#003399", "#003090", "#C8102E",
                               "#6CABDD", "#DA291C", "#241F20", "#FFF200", "#EE2737",
                               "#D71920", "#132257", "#FBEE23", "#7A263A", "#FDB913")) +
  labs(x = "\nTeam",
             y = "Average Expected Goals Conceded\n",
             colour = "",
             title = "Premier League Average xGA\n")
```


## Conclusion

- They create plenty chances to score

- They rarely concede, and have one of the best defenses in the league

- They don't have a great expected goal vs goal scored rate for a team of their calibre

- This suggests they should invest in a clinical forward


## How do you define whether a forward is 'clinical' or not?

- They score goals (lots of goals)

- They perform at a high level

- They outperform their xG

# Who are the best strikers in Europe? 

---

```{r echo=FALSE,out.height="600px",out.width="1000px"}
player_data_pl <- read_csv("man_city_transfer_project/data/england/2020.21.EPL.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_ligue_1 <- read_csv("man_city_transfer_project/data/france/2020.21.Ligue1.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_bundesliga <- read_csv("man_city_transfer_project/data/germany/2020.21.Bundesliga.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_la_liga <- read_csv("man_city_transfer_project/data/spain/2020.21.La.Liga.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_serie_a <- read_csv("man_city_transfer_project/data/italy/2020.21.SerieA.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)


full_join(player_data_bundesliga, player_data_la_liga) %>%
  full_join(player_data_ligue_1) %>%
  full_join(player_data_pl) %>%
  full_join(player_data_serie_a) %>% 
  mutate("xGvsG" = goals - xG)  %>%
  filter(goals >= 20)%>%
 arrange(desc(xGvsG)) %>%
  ggplot() +
  aes(x = reorder(player_name, -xGvsG), y = xGvsG, fill = player_name) +
  geom_col(colour = "#000000") +
  theme_economist() +
 theme(legend.position = "none", axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = c("#e1000f", "#1C63B7", "#87D8F7", "#FFFFFF", "#482E92",                                    "#FDE100","#FFE667", "#132257", "#00529F", "#004170",                                    "#A50044", "1E71B8", "#CB3524", "#36384a", "#C8102E",                                    "#DC052D","#010E80", "#701D2F", "#c8102e", "#65B32E")) +
  labs(x = "\nPlayer",
           y = "Goals minus Expected Goals\n",
            colour = "",
           title = "Goals scored vs xG\n ",
       subtitle = "Top 20 scorers in the Top 5 European Leagues\n
       (Minimum twenty goals scored)\n")
```

## Lessons for the future 

- Getting open source Football Data is very hard

- Hex codes are infuriating

- Everything will always takes longer than you think

- Don't use a presentation tool you've never used before (especially if it's ioslides)

# Questions?