---
title: "R Notebook"
output: html_notebook
---

Area(s) city are weak? 
- graph to show their xG is high
- graph to show that their xG vs G isn't great for a team of their caliber
- graph to show their defensive quality
Conclusion = They need to invest in a striker.

Best strikers in europe?
who has the xG vs G (min 20 goals)

Conclusion. They should go for one of these 5 options based on my research.
- Lewa, Kramaric, Moreno, Benzema, Haaland        

```{r}
library(tidyverse)
```

## Premier League data and how Manchester city compare to the rest

```{r}
team_xg_data <- read_csv("data/understat_per_game.csv") %>% 
  rename("conceded" = "missed")

# Average number of goals scored a match in the Premier League
  g_data_2019 <- team_xg_data %>%
 filter(league == "EPL",
        year == 2019) %>%
  group_by(team) %>%
  summarise(mean(scored))
  # Mean xG 
    xG_data_2019 <- team_xg_data %>%
 filter(league == "EPL",
        year == 2019) %>%
  group_by(team) %>%
  summarise(mean(xG))

# Expected number of goals conceded a game in the Premier League
pl_xga_data <- team_xg_data %>%
   filter(league == "EPL",
          year == 2019) %>%
  group_by(team) %>%
  summarise(mean(xGA))
# Average number of goals conceded a game in the Premier League
pl_ga_data <- team_xg_data %>%
   filter(league == "EPL",
          year == 2019) %>%
  group_by(team) %>%
  summarise(mean(conceded))

# Difference between expected goals and goals scored in the Premier League
pl_xg_data_2019 <- team_xg_data %>%
 filter(league == "EPL",
        year == 2019)%>%
  group_by(team) %>%
  summarise(mean(xG_diff))
```

This shows that city create among the most chances in the league and also have one of the
best defensive records in the league. Where they lack is in xG vs G. Therefore, I'd recommend
they target a striker with a high conversion rate.


## Striker Player Data
```{r}
player_data_pl <- read_csv("man_city_transfer_project/data/england/2020.21.EPL.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_ligue_1 <- read_csv("data/france/2020.21.Ligue1.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_bundesliga <- read_csv("data/germany/2020.21.Bundesliga.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_la_liga <- read_csv("data/spain/2020.21.La.Liga.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)

player_data_serie_a <- read_csv("data/italy/2020.21.SerieA.Meta.csv") %>%
  arrange(desc(goals)) %>%
  head(10)


top_scorers_top_5_leagues <- full_join(player_data_bundesliga, player_data_la_liga) %>%
  full_join(player_data_ligue_1) %>%
  full_join(player_data_pl) %>%
  full_join(player_data_serie_a) %>% 
  mutate("xGvsG" = goals - xG)  %>%
  filter(goals >= 20)%>%
 arrange(desc(xGvsG)) 
```

